
**Abstract:** Planespotters.net has a full database on airlines around the world and the airplanes that each owns and operates.This dataset collects the top 100+ airlines in the world (by the size of their fleet). It is combined with information found on Wikipedia on the respective airline's fleet and the average value/cost of the manufactured airplane. Dataset includes: (a)Parent Airline: i.e. International Airlines Group (IAG) (b)Airline: i.e. Iberia, Aer Lingus, British Airways…etc. which are owned by IAG (c) Aircraft Type: Manufacturer & Model (d)Current: Quantity of airplanes in Operation (e)Future: Quantity of airplanes on order, from planespotter.net (f)Order: Quantity airplanes on order, from Wikipedia (g)Unit Cost: Average unit cost of Aircraft Type, as found by Wikipedia and various google searches (h)Total Cost: Current quantity * Unit Cost (i)Average Age: Average age of "Current" airplanes by "Aircraft Type".

## The Data set:
```{r}
flight = read.csv("C:/Users/savla/OneDrive/Desktop/dataset/Fleet Data.csv")
```

**Observation** This Fleet Dataset contains 1583 rows and 11 columns with names.

## Load in Libraries:

```{r results='hide'}
library(tidyverse)
library(dplyr)
library(devtools)
library(readxl)
library(rvest)
library(sf)
library(rjson)
library(jsonlite)
library(tidycensus)
library(class)
library(stats)
library(rpart)
library(rpart.plot)
library(ggplot2)
library(readr)
library(data.table)
```


```{r}
names(flight) = c("Parent_Airline",
                  "Airline",
                  "Aircraft_Type",
                  "Current",
                  "Future",
                  "Historic",
                  "Total",
                  "Orders",
                  "Unit_Cost",
                  "Total_Cost_Current",
                  "Average_Age")
colnames<-names(flight)
```

Question 1 use str(), glipse(), dim() and summary()
```{r results='hide'}
glimpse(flight)
```
*Note:* Using glipse() function, I was able to see all rows and colums of the dataset with their names.

```{r results='hide'}
dim(flight)
```
*Note:* Using dim() function, I was able to see the number of Rows(1583) and Columns(11) 

```{r results='hide'}
str(flight)
```
*Note:* Using str() function, I was able to see the number of Rows(1583) and Columns(11) 

```{r results='hide'}
summary(flight)
```
*Note:* Using summary() function, 












